---
title: "Crocodile Behavior Tutorial"
author: "Maggie McGreal"
date: "`r Sys.Date()`"
output: html_document
---


```{r}
install.packages("googlesheets4")

```


The purpose of this R package is for analysis and visualization of remotely transmitted movement and behavioral data from animal-borne devices.... (behavior classification, )
This package will work with data obtained from custom-built satellite trackers developed by Dr. Youngho Shin
This data includes measurements obtained from sensors such as: IMU, GPS, real-time clock, depth recorder, etc.

```{r}
library(googlesheets4)
googlesheets4::gs4_deauth()
crocodiles <- read_sheet("https://docs.google.com/spreadsheets/d/1b4j42SbcN0GZR6KgHRtEmRe5-D0qzYM1mZWo9-KC4d4/edit?usp=sharing")

```


```{r}
#Install the package from git hub

devtools::install_github("mcgrealmaggie/crocodiles.pkg")


```

Function 1: sensorData
```{r}
#data <- sensorData("crocodile.csv")
```

Function 2: createDatetime
```{r}
#data <- createDatetime(data, "date", "time")
```

Function 3: calculateODBA
```{r}
#data <- calculateODBA(data = data, X, Y, Z)
```

Function 4: performLM
```{r}
#lm <- performLM(data, "ODBA", "behavior")
```

Function 5: plotModel
```{r}
#plotModel(data, "behavior", "ODBA")
```

Function 6: mapODBA
```{r}
#mapODBA(data, "long", "lat", "ODBA")
```

Function 7: mapBehavior
```{r}
#mapBehavior(data, "long", "lat", "behavior")
```
